The investment strategy based on behavior of artificial earthworm for use in algotrading

### 2014-04-08 Article/Article20140406Granada/LatexArt/article_main.tex
Bibliography 1st version

Algotrading is rapidly growing new form of trading on financial markets where investor (that makes decisions) is replaced with automaton (computer program) []. Such machine can perform operations (opening and closing positions, placing orders to open such positions that meet certain conditions) directly in an automated brokerage platform environment. It may also be a recommendation system that helps investor with making correct decisions.\\

Such computer applications (bots) require carefully prepared and tested software that most often needs to adapt to the changing situations on the observed markets. Modern trends, in the development of this software, include sophisticated methods of: artificial intelligence, pattern recognition, classification and machine learning. Those methods are interpenetrating and complementary and are also expanded with information based on fundamental analysis, for example, automatically extracted from the Internet \cite{elder} [Schwager, Elder, … nasze hindawi].\\

@@

Let's consider an artificial worm, which behaves like its natural inspirer. Let this artificial worm have a length of p1 candles (vectors representing the knowledge about events occurring within the contractual part of the time series). Let her movement come down to the successive ,,eating'' one candle and simultaneous ,,discarding'' of one candle --- always having p1 candles. Thus the earthworm moves incrementally at length of time called the period of the candle. I-th candle is  defined as a row of data of values  $Oi$, $Hi$, $Li$, $Ci$ (Open, High, Low and Close). These values represent (in order): the value at the beginning of this period, the maximum value within the period, the minimum and the closing values - at the end of the candle.\\
Let's assume that candle can have color (often in a number of brokerage platforms, this procedure is used). Therefore candles may be red (when the opening of the candle is higher than its closing $Oi>Ci$) and green (when the situation is opposite $Oi <Ci$).\\
Let's also assume --- it is a primary strategy rule - that when an artificial earthworm has a significant advantage of red candles, a short position is opened or continued. When earthworm has the significant advantage of green candles, a long position is opened or continued. Imprecise term ,,significant'' is defined by the use of the second parameter p2 according to the rules:
\begin{equation}
\text{If }  Lr > p1/2 -p2 \text{ then short;}
\end{equation}
opening of short position will be suggested;
\begin{equation}
\text{If }  Lg > p1/2 -p2 \text{ then long;}
\end{equation}
opening of long position will be suggested;
\begin{equation}
\text{If }  Lg < p1/2 -p2 \text{ and } Lr< p1/2-p2 \text{ then none;}
\label{eq:eq3}
\end{equation}
the continuation of state will be suggested (position or no opened position).\\

@@

From these results it can be concluded that the strategy is working with a short shelf life characteristics. The best results are achieved in periods of about 10-20 candles, it is unlike any other strategy presented in [Hindawi] where the best results were achieved for the test periods with a length of 80-100 candles.\\
All the above presented results refer to simulations in Matlab.

@@

Presented in the paper the concept of an innovative investment strategy gave satisfactory results for a certain pair as the result of periods of machine learning. For example, in the studied data for two key market currency pairs sampled every hour , the best value is 80 hours of teaching and 15 hours of testing. One can assume that this relationship may change in a different environment (for other markets, for other periods) by becoming a sort of next parameter that optimal values should be sought. It should be emphasized that the study was conducted on data not used in the learning phase and for the basic principle of machine learning achieved results that would have satisfied many a practice. It should also be noted that outside research conducted in environment completely specified (for constant unchanging data stored in Matlab ) was also performed simulations in terms of actual brokerage platform. The results obtained in both environments are similar in the sense of direction changes and logic behavior caused by the user received by the algorithm implemented in the Matlab and Metatrader. Differences explained in this work .
The direction of further research will test the hypothesis that the variability periods to test the learner can improve the quality of prediction. This hypothesis seems to be interesting against the background of previous experiences and results, eg for different studied strategies, including [ Hindawi ] obtained surprisingly different optimal length of periods of learning and testing.

>>>>

Algotrading is rapidly growing new form of trading on financial markets where investor (that makes decisions) is replaced with automaton (computer program) \cite{Leshik2011}. Such machine can perform operations (opening and closing positions, placing orders to open such positions that meet certain conditions) directly in an automated brokerage platform environment. It may also be a recommendation system that helps investor with making correct decisions.\\

Such computer applications (bots) require carefully prepared and tested software that most often needs to adapt to the changing situations on the observed markets. Modern trends, in the development of this software, include sophisticated methods of: artificial intelligence, pattern recognition, classification and machine learning. Those methods are interpenetrating and complementary and are also expanded with information based on fundamental analysis, for example, automatically extracted from the Internet \cite{elder}\cite{Wilinski2014}\cite{Schwager1996}.\\

@@

Let's consider an artificial worm, which behaves like its natural inspirer. Let this artificial worm have a length of p1 candles (vectors representing the knowledge about events occurring within the contractual part of the time series). Let her movement come down to the successive ,,eating'' one candle and simultaneous ,,discarding'' of one candle --- always having p1 candles. Thus the earthworm moves incrementally at length of time called the period of the candle. I-th candle is  defined as a row of data of values  $Oi$, $Hi$, $Li$, $Ci$ (Open, High, Low and Close). These values represent (in order): the value at the beginning of this period, the maximum value within the period, the minimum and the closing values - at the end of the candle.\\
Let's assume that candle can have color (often in a number of brokerage platforms, this procedure is used). Therefore candles may be red (when the opening of the candle is higher than its closing $Oi>Ci$) and green (when the situation is opposite $Oi<Ci$).\\
Let's also assume --- it is a primary strategy rule - that when an artificial earthworm has a significant advantage of red candles, a short position is opened or continued. When earthworm has the significant advantage of green candles, a long position is opened or continued. Imprecise term ,,significant'' is defined by the use of the second parameter p2 according to the rules:
\begin{equation}
\text{If }  Lr > p1/2 -p2 \text{ then short;}
\end{equation}
opening of short position will be suggested;
\begin{equation}
\text{If }  Lg > p1/2 -p2 \text{ then long;}
\end{equation}
opening of long position will be suggested;
\begin{equation}
\text{If }  Lg < p1/2 -p2 \text{ and } Lr< p1/2-p2 \text{ then none;}
\label{eq:eq3}
\end{equation}
the continuation of state will be suggested (position or no opened position).\\

@@

From these results it can be concluded that the strategy is working with a short shelf life characteristics. The best results are achieved in periods of about 10-20 candles, it is unlike any other strategy presented in \cite{Wilinski2014} where the best results were achieved for the test periods with a length of 80-100 candles.\\
All the above presented results refer to simulations in Matlab.

@@

Presented in the paper the concept of an innovative investment strategy gave satisfactory results for a certain pair as the result of periods of machine learning. For example, in the studied data for two key market currency pairs sampled every hour , the best value is 80 hours of teaching and 15 hours of testing. One can assume that this relationship may change in a different environment (for other markets, for other periods) by becoming a sort of next parameter that optimal values should be sought. It should be emphasized that the study was conducted on data not used in the learning phase and for the basic principle of machine learning achieved results that would have satisfied many a practice. It should also be noted that outside research conducted in environment completely specified (for constant unchanging data stored in Matlab ) was also performed simulations in terms of actual brokerage platform. The results obtained in both environments are similar in the sense of direction changes and logic behavior caused by the user received by the algorithm implemented in the Matlab and Metatrader. Differences explained in this work .
The direction of further research will test the hypothesis that the variability periods to test the learner can improve the quality of prediction. This hypothesis seems to be interesting against the background of previous experiences and results, eg for different studied strategies, including \cite{Wilinski2014} obtained surprisingly different optimal length of periods of learning and testing.
